\section{Related Work}

\paragraph{Routability-Aware Objectives.}
Routability-aware placement exposes downstream routing or timing cost to an
analytical optimizer. RePlAce established the electrostatic density
formulation~\cite{replace}, and DREAMPlace made analytical placement
GPU-efficient in PyTorch~\cite{dreamplace}. One line adds expert symbolic
terms: RUDY estimates routing demand from placement
geometry~\cite{rudy}. A second line inserts learned downstream surrogates.
LaMPlace, GOALPlace, and RoutePlacer connect placement to post-route timing,
design-rule checking (DRC), or congestion through masks, graph neural
networks, and learned routability signals~\cite{lamplace,goalplace,routeplacer}.
These methods add downstream awareness through fixed expert terms or learned
components. CoEvoP\&R searches for readable symbolic objective forms that
still execute inside the analytical placer.

\paragraph{Foundation-Model Placement Search.}
Placement search methods differ by the artifact they modify. Reinforcement
learning places macros sequentially~\cite{mirhoseini2021}. AutoDMP tunes
macro-placement parameters with multi-objective Bayesian
optimization~\cite{autodmp}, and HyperPlace uses an off-the-shelf LLM to tune
GPU-placer hyperparameters~\cite{hyperplace}. EvoPlace evolves optimizer
code, including initialization, preconditioning, and update components, with
HPWL feedback~\cite{evoplace}. VeoPlace uses a vision-language model (VLM) to
guide macro-placement actions through evolutionary search~\cite{veoplace}.
ChiPBench provides end-to-end validation for whether
placement gains survive downstream flow~\cite{chipbench}. These systems
search parameters, optimizer components, or macro actions around the placement
objective. CoEvoP\&R targets the differentiable objective itself and uses
scheduled routed promotions to condition later objective proposals.

\paragraph{Executable LLMs for EDA.}
LLM-based EDA work often uses models for text, register-transfer level (RTL),
or prediction tasks. ChipNeMo adapts foundation models for chip-design
tasks~\cite{chipnemo}. RTLCoder generates Verilog from natural
language~\cite{rtlcoder}. MasterRTL and VeriLoC predict pre-synthesis power,
performance, area, or line-level design quality from
Verilog~\cite{masterrtl,veriloc}. A second line turns LLM output into
executable programs. FunSearch searches over Python programs for
combinatorial mathematics~\cite{funsearch}. Eureka evolves reward functions
from training-statistics feedback~\cite{eureka}. Evolution of Heuristics
(EoH) co-evolves natural-language thoughts and executable heuristic
code~\cite{eoh}. LLM-SR recovers symbolic scientific equations under
automated evaluation~\cite{llmsr}. AlphaEvolve extends the pattern to full
code files across mathematics and computing infrastructure~\cite{alphaevolve},
and OpenEvolve provides a public implementation~\cite{openevolve}. These
systems evaluate generated programs after execution. CoEvoP\&R places
generated code inside DREAMPlace's autograd path, so validity requires finite
objective values, finite gradients, and optimizer-compatible terms before
routed validation.
